---
title: "Exploratory Data Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    theme: flatly
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tidyr)
library(naniar)
library(stringr)
library(janitor)
library(plotly)
library(factoextra)
library(gridExtra)
library(kableExtra)
library(knitr)
library(tableone)
library(fmsb)
```

Now, we get our neat dataset, `superheroes`. For this dataset, we get boxplots for the numerical variables `intelligence`, `strength`, `speed`, `durability`, `power`, `combat`, `height_cm`, `weight_kg` which helps us to understand their distribution. For the categorical variables `publisher`, `alignment`, `gender`, `eye_color`, `hair_color`, we also analyzed them accordingly. At the end of this section, we performed a clustering analysis of all heroes without missing data and categorized them into three groups: the balanced group, the strength group, and the mediocre group.

## Boxplot and correlation matrix for all numeric variables

```{r message = FALSE, warning = FALSE}
superheros <- read_csv("data/superheroes.csv")

# Box plot

numeric_vars <- superheros |>
  select(intelligence, strength, speed, durability, power, combat, height_cm, weight_kg)

plots <- lapply(names(numeric_vars), function(var) {
  plot_ly(y = numeric_vars[[var]], type = "box", name = var)
})

boxplot_combined <- subplot(plots, nrows = 3, margin = 0.05)

boxplot_combined


# Correlation Matrix

cor_matrix <- cor(numeric_vars, use = "complete.obs")

cor_heatmap <- plot_ly(
  z = cor_matrix,
  x = colnames(cor_matrix),
  y = colnames(cor_matrix),
  type = "heatmap",
  colors = colorRampPalette(c("lightblue", "blue", "darkblue"))(100)
) %>%
  layout(
    title = "Correlation Matrix Heatmap",
    xaxis = list(title = "Variables"),
    yaxis = list(title = "Variables")
  )

cor_heatmap
```

The first figure shows the distribution of different variables using box plots to visualize the statistical characteristics of multiple superhero attributes.

`Intelligence`: The median is around 50, with most values concentrated between 25 and 75, and some low outliers.

`Strength`: The distribution is fairly even, with almost no outliers. The median is approximately 50, and the range spans from 0 to 100.

`Speed`: Most data points are concentrated in the middle range, with a median slightly above 50, and no obvious outliers.

`Durability`: The median is around 50, and the overall distribution is even, ranging from 0 to 100, with almost no outliers.

`Power`: The median is close to 50, with a wide range of values. The overall distribution is relatively symmetric.

`Combat`: The median is near 50, and the data has a wide range, with an even distribution and no clear outliers.

`Height (cm)`: Height shows considerable variation, with some significant outliers, particularly above 500 cm, indicating some extremely tall superheroes. The median is below 200 cm, with most data points below 500 cm.

`Weight (kg)`: The median weight is around 100 kg. There are some very large outliers (over 400 kg), with most values concentrated below 200 kg.

The second heatmap represents the correlation matrix between several superhero attributes, including intelligence, strength, speed, durability, power, combat, height (in cm), and weight (in kg). The color gradient ranges from light blue to dark blue, where darker shades indicate a stronger positive correlation between two variables. There is a strong correlation between strength and durability (0.65). This suggests that superheroes who are strong also tend to be durable. Height and weight also show a very high correlation (0.69), indicating that taller superheroes tend to be heavier.

## Bar chart of publisher

```{r warning=FALSE}
publisher_counts <- superheros |>
  count(publisher) |>
  arrange(desc(n))

publisher_bar <- plot_ly(publisher_counts, x = ~reorder(publisher, -n), y = ~n, type = 'bar', marker = list(color = '#aec7e8')) |>
  layout(title = "Publisher Distribution (Bar Chart - Sorted by Count)",
         xaxis = list(title = "Publisher"),
         yaxis = list(title = "Count"))

publisher_bar
```

This bar chart represents the distribution of superheroes by publisher, sorted by the number of heroes. The chart shows that Marvel Comics and DC Comics are by far the most dominant publishers, with Marvel having the highest count at 339 heroes, followed by DC Comics with 188 heroes. 

## Pie Chart of alignment

```{r}
alignment_counts <- superheros |>
  count(alignment)

alignment_pie <- plot_ly(alignment_counts, labels = ~alignment, values = ~n, type = 'pie',
                        marker = list(colors = c('#98df8a', '#ffbb78', '#9edae5', '#f7b6d2'))) |>
  layout(title = "Alignment Distribution",
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

alignment_pie
```

This pie chart illustrates the distribution of superhero alignments. The majority of superheroes, about 67.6%, fall under the "Good" category, represented in orange. The second largest segment is the "Bad" alignment, accounting for 28.2%, shown in green. The "Neutral" category, depicted in light blue, represents 3.28% of the characters. Lastly, a tiny slice, labeled as "null" in pink, makes up 0.958% of the characters.

## Further analysis for Marvel and DC

```{r}
marvel_alignment <- superheros |>
  filter(publisher == "Marvel Comics") |>
  count(alignment)

dc_alignment <- superheros |>
  filter(publisher == "Dc Comics") |>
  count(alignment)

# Plot alignment distribution for Marvel Comics
marvel_plot <- plot_ly(marvel_alignment, labels = ~alignment, values = ~n, type = 'pie',
                      marker = list(colors = c('#aec7e8', '#ffbb78', '#98df8a', '#f4cccc'))) |>
  layout(title = "Alignment Distribution - Marvel Comics",
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

# Plot alignment distribution for DC Comics
dc_plot <- plot_ly(dc_alignment, labels = ~alignment, values = ~n, type = 'pie',
                   marker = list(colors = c('#c5b0d5', '#c49c94', '#f9cb9c', '#d9d9d9'))) |>
  layout(title = "Alignment Distribution - DC Comics",
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

marvel_plot
dc_plot
```

Since Marvel and DC are the most popular publishers, we checked the distributions of superhero alignments within these two publishers. The two pie chats show the similar distribution as the overall alignments distribution.

## Bar Chart for eye and hair color

```{r warning=FALSE}
eye_color_counts <- superheros |>
  count(eye_color)

eye_color_bar <- plot_ly(eye_color_counts, 
                         x = ~n, 
                         y = ~reorder(eye_color, n), 
                         type = 'bar', 
                         orientation = 'h', 
                         marker = list(color = "#dbdb8d")) |>
  layout(title = "Eye Color Distribution",
         xaxis = list(title = "Count"),
         yaxis = list(title = "Eye Color"))

hair_color_counts <- superheros |>
  count(hair_color)

hair_color_bar <- plot_ly(hair_color_counts, 
                          x = ~n, 
                          y = ~reorder(hair_color, n), 
                          type = 'bar', 
                          orientation = 'h', 
                          marker = list(color = "#c7c7c7")) |>
  layout(title = "Hair Color Distribution",
         xaxis = list(title = "Count"),
         yaxis = list(title = "Hair Color"))

eye_color_bar

hair_color_bar
```

This figure contains two bar charts displaying the distribution of superhero eye colors and hair colors. Overall, blue eyes and black hair are the most common features among superheroes.

## Demographic Table classified by alignment and gender

```{r message = FALSE}
superheros1 <- superheros |>
  filter(!is.na(alignment), !is.na(gender))

vars <- c("intelligence", "strength", "speed", "durability", "power", "combat", "height_cm", "weight_kg")

alignment_summary <- CreateTableOne(vars = vars, strata = "alignment", data = superheros1, test = TRUE)

alignment_summary_df <- print(alignment_summary, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)

alignment_summary_df |> as.data.frame() |>
  rename(`p-value` = p) |>
  mutate(test = "F-test") |>
  kable()

gender_summary <- CreateTableOne(vars = vars, strata = "gender", data = superheros1, test = TRUE)

gender_summary_df <- print(gender_summary, quote = FALSE, noSpaces = TRUE, printToggle = FALSE)

gender_summary_df |> as.data.frame() |>
  rename(`p-value` = p) |>
  mutate(test = "T-test") |>
  kable()

```


This table summarizes the demographic characteristics of superheroes by alignment ("Bad," "Good," "Neutral") and gender ("Female," "Male"). The mean and standard deviation are provided for attributes such as intelligence, strength, speed, and others. Significant differences are observed among alignment categories for attributes like intelligence, strength, durability, height, and weight, as indicated by F-test p-values. In the gender comparison, significant differences are found in strength, durability, height, and weight, with males generally showing higher values. These statistical tests highlight notable variations in superhero traits based on alignment and gender.

## Clustering

```{r}
numeric_vars <- superheros |>
  select(intelligence, strength, speed, durability, power, combat, height_cm, weight_kg)
superheros_clean <- superheros[complete.cases(numeric_vars), ]
numeric_vars_clean <- superheros_clean |>
  select(intelligence, strength, speed, durability, power, combat, height_cm, weight_kg)

numeric_vars_scaled <- scale(numeric_vars_clean)

set.seed(123)

kmeans_result_2 <- kmeans(numeric_vars_scaled, centers = 2, nstart = 25)
plot_2 <- fviz_cluster(kmeans_result_2, data = numeric_vars_scaled,
                       geom = "point",
                       ellipse.type = "convex",
                       ggtheme = theme_minimal()) +
  labs(title = "k = 2")

kmeans_result_3 <- kmeans(numeric_vars_scaled, centers = 3, nstart = 25)
plot_3 <- fviz_cluster(kmeans_result_3, data = numeric_vars_scaled,
                       geom = "point",
                       ellipse.type = "convex",
                       ggtheme = theme_minimal()) +
  labs(title = "k = 3")

kmeans_result_4 <- kmeans(numeric_vars_scaled, centers = 4, nstart = 25)
plot_4 <- fviz_cluster(kmeans_result_4, data = numeric_vars_scaled,
                       geom = "point",
                       ellipse.type = "convex",
                       ggtheme = theme_minimal()) +
  labs(title = "k = 4")

grid.arrange(plot_2, plot_3, plot_4, ncol = 3)
```

### List of clustering

This is the list of superheroes in each clustering group. Let's see where your superhero is!

```{r}
superheros_clean$cluster <- kmeans_result_3$cluster

superheros_clean |>
  group_by(cluster) |>
  summarise(heroes = paste(name, collapse = ", ")) |>
  kable()
```

### Check capability statistics within each cluster group

```{r}
superheros_clean |>
  group_by(cluster) |>
  summarise(
    intelligence_mean = mean(intelligence, na.rm = TRUE),
    intelligence_sd = sd(intelligence, na.rm = TRUE),
    strength_mean = mean(strength, na.rm = TRUE),
    strength_sd = sd(strength, na.rm = TRUE),
    speed_mean = mean(speed, na.rm = TRUE),
    speed_sd = sd(speed, na.rm = TRUE),
    durability_mean = mean(durability, na.rm = TRUE),
    durability_sd = sd(durability, na.rm = TRUE),
    power_mean = mean(power, na.rm = TRUE),
    power_sd = sd(power, na.rm = TRUE),
    combat_mean = mean(combat, na.rm = TRUE),
    combat_sd = sd(combat, na.rm = TRUE),
    height_mean = mean(height_cm, na.rm = TRUE),
    height_sd = sd(height_cm, na.rm = TRUE),
    weight_mean = mean(weight_kg, na.rm = TRUE),
    weight_sd = sd(weight_kg, na.rm = TRUE)
  ) |>
  pivot_longer(-cluster, names_to = c("variable", ".value"), names_sep = "_") |>
  mutate(mean_sd = paste0(round(mean, 2), " (", round(sd, 2), ")")) |>
  select(variable, cluster, mean_sd) |>
  pivot_wider(names_from = cluster, values_from = mean_sd, names_prefix = "cluster") |>
  kable()
```

### Radar Plot!

```{r warning=FALSE}
cluster_averages <- superheros_clean |>
  group_by(cluster) |>
  summarise(
    intelligence = mean(intelligence, na.rm = TRUE),
    strength = mean(strength, na.rm = TRUE),
    speed = mean(speed, na.rm = TRUE),
    durability = mean(durability, na.rm = TRUE),
    power = mean(power, na.rm = TRUE),
    combat = mean(combat, na.rm = TRUE)
  )

radar_data <- as.data.frame(cluster_averages)
radar_data <- radar_data[, -1]

radar_data <- rbind(rep(100, 6), rep(0, 6), radar_data)

colors_border <- c('#f08080', '#90ee90', '#87cefa')
colors_in <- c('#ffcccb', '#c6e7c6', '#b0e0e6')

radarchart(radar_data, axistype = 2, 
           pcol = colors_border, 
           pfcol = adjustcolor(colors_in, alpha.f = 0.6), plwd = 2.5, plty = 1,
           cglcol = "grey", cglty = 1, 
           axislabcol = "black")

legend(x = "topright", legend = paste("Cluster", 1:3), pch = 20, col = colors_border, text.col = "black", cex = 1.2, pt.cex = 2)
```